---
title: "Working with Markdown. My first R markdown document."
author: "Daniel Penados-Richter"
date: "2024-01-23"
output: 
  html_document: 
    theme: journal
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For Markdown, use asteriks to *indicate italic text*. 

Double line return for a new chunk of text like this. 
if not is like this. 

You could always try this:

Name  
University of Vermont

# Head line are made with hash marks

## Two head marks make it smaller

### Smaller

#### So one...


**Bold face this way**

>Offset quote which embed and looks really nice when you have extended material that you would like to see. 

back to plain text


to find more useful code go to help>markdown quick reference

List  
*Item 1  
*Item 2  
  - Spacing incidents
  
  [Set up a link with text](website_adress)
  
  also footnotes^[whatever foot note]
  
We can create a fencing for code: 

```
fencing serts boundaries and everthin inside the fencing *is shown as literal text* without rendering

```
The Rmarkdown code help us Latex code becouse is excelent for any mathematical lenguaje. 

to enter use a double $ signe

$$
a+b=c
$$

There is a second way to enter to latex lenguaje: $a+b=c$ like this  
Now the material is ibedeed in my sentence. 

A couple of examples in LateX: 

The Curly brackets set that the particular code before apply to everything between it:  

$$ H_0 = Z_{a + b} $$

## we can also do superscripts in LaTeX

$$ A = cA^z$$

$$S=cA^z-1 + Z_{2 + x}$$
## Fraction 

$$\alpha = \frac{\beta}{\delta + \gamma_x}$$

# We can do sumatorias. 

$$z - \sum_{i=1}^X{K}$$


What happens is a command is a symbols that we actually want to use? Well u use a backslash to scape the code. 


$$\backslash \alpha \le b \backslash$$
 To introduce plain text we can use a M box. 

$$P(\mbox{Ocurrence Of Species A}) = Z$$


### You can put things in line with the text. 


But remember this does not have to be as an equation. We could also embed them in the sentnces with a single dollar sign. $\pi + 3$


We can also use markdone fencing in file with a single backtick like this: `pi + 3` but remmeber 

```3 backtick creats an isolated chunk of text```


We can even execute R code inside of markdown. With r between backtick. Everithing after it is gonna be interpreted as R code `r pi +3`. In this case pi is store as the number 3.14.... and now is the results of the calculation explain before.  


This come that we can do R code. Real chunks of R code. If you do 3 back ticks and {R} well creat a R code that will run as a code. 

```{r}
Pred <- seq(1, 10) #Make a vercotr of integer fron 1 to 10. 
Res <- runif(10)   #generate 10 draws from a randon uniform (0,1) distristribution

#print the random numbers
print(Res)

```


As you see above,  It print the code, run it and print the result. it and printed 10 vrandom number between 0 and 1. Also can print the actual final graph, or whatever you want that R does. 

we can ad an instruction to print the code, but do not print the output. 

```{r eval=FALSE}
Pred <- seq(1, 10) #Make a vercotr of integer fron 1 to 10. 
Res <- runif(10)   #generate 10 draws from a randon uniform (0,1) distristribution

#print the random numbers
print(Res)

```

This other will print the result but will not show the code

```{r eval=TRUE, echo=FALSE}
Pred <- seq(1, 10) #Make a vercotr of integer fron 1 to 10. 
Res <- runif(10)   #generate 10 draws from a randon uniform (0,1) distristribution

#print the random numbers
print(Res)

```


```{r}
Pred <- seq(1, 10) #Make a vercotr of integer fron 1 to 10. 
Res <- runif(10)   #generate 10 draws from a randon uniform (0,1) distristribution

#print the random numbers
print(Res)

```



##Speaking a little bit about and R script: 

<cntrl><return> runs code by code. and what is the difference with <command><return>
