---
title: "Homework12. Ggplot. "
author: "Daniel Penados-Richter"
date: "2024-04-22"
output: html_document
---

```{r setup, include=FALSE}
library(colorspace)
library(ggplot2)
library(ggthemes)
library(patchwork)
library(cowplot)
library(wesanderson)
library(colorblindr)
library(ggsci) 
library(scales)
library(patchwork)
library(tidyverse)
library(ggpubr)
```

Lets star by importing our data. 

```{r}
#Call the data
abl<-read.csv("Abl.csv", sep = ",", header = TRUE)
glimpse(abl) #Its already in the long form for ggplot

#Lets create a simple box plot, 
p1<-ggplot(abl, aes(x = Grupes, y = Value, fill = Treatment)) +
  geom_boxplot()

print(p1) 

```

This is not very pretty. I do not like the color, my name groups are awful, no points or sd in graph. Do not like the backgroun and other stuff. So lets change it. 

Lets do a couple of things firts: 
1. Rearrange the other of the groups to have consistancy accross figures and other results. 
2. Change color, crating a variable name fill_colors, that way we could use the same ones across a paper. 
3. change the theme a use it consistently across the paper. 
4. Lastly add labels. 

```{r}
 #Define the desired fill color
fill_colors <- c("DMSO" = "thistle", "STI571" = "chocolate")

p1 <- ggplot(abl, aes(x = factor(Grupes, levels = c("DCBLD2", "PDGFRB", "A_WT", "A_D842V")), y = Value, fill = Treatment)) + geom_boxplot() + #Notice the groups order is define. 
  scale_fill_manual(values = fill_colors) + #Fill accordingly to groups 
  labs( x = "Groups", #Labels
       y = "pY Signal (Fold to base)",
       fill = "Treatment") +
  theme_light()
# Display the plot
print(p1)

```

This is starting to look way better. But there is a couple of extra things to do. 
1. Change the names to of goup variables. Notice we are using a short version to work in R, but this can be change. 

```{r}
# Change the group names
new_names <- c("Control", "PDGFRß", "PDGFRA WT", "PDGFRA D842V")

# Updated plot with new group names
p2 <- p1 + scale_x_discrete(labels = new_names)
print(p2)
```
The last this that is missing to be perfect is to the results of out post hoc test. We are going to perform an anova, comparing each group within treatment. 

```{r}

p2 <- ggplot(abl, aes(x = factor(Grupes, levels = c("DCBLD2", "PDGFRB", "A_WT", "A_D842V")), y = Value, fill = Treatment)) + 
  geom_boxplot() + 
  scale_fill_manual(values = fill_colors) + 
  labs(x = "Groups", y = "pY Signal (Fold to base)", fill = "Treatment") +
  theme_light() +
  stat_compare_means(aes(group = Treatment), method = "t.test", label = "p.signif") #Use this function to imput the post hoc test values 

p3 <- p2 + scale_x_discrete(labels = new_names)
print(p3)
```

Now lets extract the figure to a new pdf, jpg or tiff. 

```{r}
#Save to PDF
ggsave("plot.pdf", p3, device = "pdf")

#Save o JPG
ggsave("plot.jpg", p3, device = "jpg")

#Lastly, save tiff 300 dpi as usually they ask. 

ggsave("plot.tiff", p3, device = "tiff", dpi = 300)

```
