---
title: "Homework4. Atomic Vectors"
author: "Daniel Penados-Richter"
date: "2024-02-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable z and print the value stored in z.

```{r}
x <- 1.1 
a <- 2.2 
b <- 3.3

z<-(x^(a^b))
print(z)

z<-(x^a)^b
print(z)

z<-(3*x^3)+(2*x^2) + 1
print(z)


```

Using the rep and seq functions, create the following vectors:

(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)


```{r}
a<-seq(from =1, to = 8)
b<-seq(from = 7, to =1)
z<-c(a,b)
print(z)

```

(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)

```{r}
a<-seq(from=1, to= 5)
z<-rep(a, time=a)
print(z)


```

(5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)

```{r}

a<-seq(from=1, to= 5)
b<-rev(a)
z<-rep(b, time=a)
print(z)



```



Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions ```(asin(), acos(), or atan())```, convert these numbers into polar coordinates (If you don’t know what polar coordinates are, read about them on the web here, here, or in your calculus textbook).


```{r}
x<-100
y<-2500
theta<-atan2(x, y)
r<-sqrt((x^2)+(y^2))
Polar<-c(r, theta)
print(Polar)

```

Create a vector ```queue <- c("sheep", "fox", "owl", "ant")``` where queue represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update queue as:

the serpent arrives and gets in line;

the sheep enters the ark;

the donkey arrives and talks his way to the front of the line;

the serpent gets impatient and leaves;

the owl gets bored and leaves;

the aphid arrives and the ant invites him to cut in line.

Finally, determine the position of the aphid in the line.


```{r}
queue <- c("sheep", "fox", "owl", "ant")
print(queue)
queue<-c(queue, "serpent")
print(queue)
queue<-c(queue[2:5])
print(queue)
queue<-c("donkey", queue)
print(queue)
queue<-c(queue[1:4])
print(queue)
queue<-c(queue[1:2], queue[4])
print(queue)
queue<-c(queue[1:2], "aphid",queue[3])
print(queue)
print(which("aphid" == queue))
```


Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7. You will need one of the arithmetic operators on this cheat sheet.


```{R}

a<-1:100
b<- which(a%%2!=0 & a%%3!=0 & a%%7!=0)
print(b)



```

