---
title: "Homework7. Testing Hyphotesis"
author: "Daniel Penados-Richter"
date: "2024-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

 

### Lets start practicing an anova with a random data generator. For 4 groups. 

```{r}
nGroup <- 4 # number of treatment groups
nName <- c("Cntrl","Treat1", "Treat2", "Treat3") # names of groups
nSize <- c(12,17,9,12) # number of observations in each group
nMean <- c(40,41,60,59) # mean of each group
nSD <- c(5,5,5,5) # standardd deviation of each group

ID <- 1:(sum(nSize)) # id vector for each row
resVar <- c(rnorm(n=nSize[1],mean=nMean[1],sd=nSD[1]),
            rnorm(n=nSize[2],mean=nMean[2],sd=nSD[2]),
            rnorm(n=nSize[3],mean=nMean[3],sd=nSD[3]),
            rnorm(n=nSize[4],mean=nMean[4],sd=nSD[4]))
Groups <- rep(nName,nSize)
DataFrame <- data.frame(ID,Groups,resVar)
str(DataFrame)
head(DataFrame)
```

#### As you can see we have 4 groups so we will evalute the difference between groups using an ANOVA.

```{r}

aov.data <- aov(resVar~Groups,data=DataFrame)
print(aov.data)
print(summary(aov.data))

```

### We have a according to the Pr(>F) (p value) there is a significant difference among groups. 

## Now lets graph it!

```{r}
my_plot <- ggplot(data=DataFrame,aes(x=Groups,y=resVar))+
           geom_boxplot()
print(my_plot)

```
Loking at the graph. We are pretty confidence that there is a significant difference among groups. Therefore is possible to hyphotesis that the different groups are Treat2 and Treat3 compare to control and Treat1. But we will need to perform a post-Hoc test to adress this. 


## What happen if we derecese the sample size? All variable will have an n=3

```{r, echo=FALSE}

nSize <- c(3,3,3,3) # number of observations in each group
nMean <- c(40,41,60,59) # mean of each group
nSD <- c(5,5,5,5) # standardd deviation of each group

ID <- 1:(sum(nSize)) # id vector for each row
resVar <- c(rnorm(n=nSize[1],mean=nMean[1],sd=nSD[1]),
            rnorm(n=nSize[2],mean=nMean[2],sd=nSD[2]),
            rnorm(n=nSize[3],mean=nMean[3],sd=nSD[3]),
            rnorm(n=nSize[4],mean=nMean[4],sd=nSD[4]))
Groups <- rep(nName,nSize)
DataFrame <- data.frame(ID,Groups,resVar)

aov.data <- aov(resVar~Groups,data=DataFrame)
print(summary(aov.data))

```

#### There is still a significant difference. Mainnly becouse the means are very different. But waht if they are not? 

```{r, echo=FALSE}
nMean <- c(40,41,45,46) # mean of each group
nSD <- c(5,5,5,5) # standardd deviation of each group

ID <- 1:(sum(nSize)) # id vector for each row
resVar <- c(rnorm(n=nSize[1],mean=nMean[1],sd=nSD[1]),
            rnorm(n=nSize[2],mean=nMean[2],sd=nSD[2]),
            rnorm(n=nSize[3],mean=nMean[3],sd=nSD[3]),
            rnorm(n=nSize[4],mean=nMean[4],sd=nSD[4]))
Groups <- rep(nName,nSize)
DataFrame <- data.frame(ID,Groups,resVar)

aov.data <- aov(resVar~Groups,data=DataFrame)
print(summary(aov.data))

```
Now there is no Significant difference. Lets try to increase the sample size again, keeping the same means. 


```{r, echo=FALSE}

nSize <- c(50, 54, 60, 57) # number of observations in each group
nMean <- c(40,41,60,59) # mean of each group
nSD <- c(5,5,5,5) # standardd deviation of each group

ID <- 1:(sum(nSize)) # id vector for each row
resVar <- c(rnorm(n=nSize[1],mean=nMean[1],sd=nSD[1]),
            rnorm(n=nSize[2],mean=nMean[2],sd=nSD[2]),
            rnorm(n=nSize[3],mean=nMean[3],sd=nSD[3]),
            rnorm(n=nSize[4],mean=nMean[4],sd=nSD[4]))
Groups <- rep(nName,nSize)
DataFrame <- data.frame(ID,Groups,resVar)

aov.data <- aov(resVar~Groups,data=DataFrame)
print(summary(aov.data))

```
### We have a significance difference again. Meaning that when the means are very different a small sample size will be enough to evidence the difference. But the smaller the difference, you will need to increase the sample size, to actually evidence a difference. 

#### Side Note: Be careful, this does not mean that just becouse we increase our sample size, we are always going to evidence a sigfificant difference. 


## Now lets try the same but with real data. 
### Testing hyphtesis with my data. Were we evaluate the importance of seven intracellular tyrosine of DCBLD2, for its phposphorylation driven by PDGFR. 

### For this we have two group. DCBLD2 WT and Mutant DCBLD Y7F (Seven intracellular Y mutated to F).

```{r}
my_data<-read.table("Homework7.csv", header = TRUE, sep = ",")
head(my_data)
```
As you can see, I have 2 variable, therefore I will performe a t test to evaluate if there is a significant difference between the two groups. 

```{r}
T.data <- t.test(pyfold~Treatment,data=my_data)
print(T.data)
print(summary(aov.data))
```
#### There is a significant difference between groups. Lets graph this. 

```{r}
my_plot <- ggplot(data=my_data,aes(x=Treatment,y=pyfold))+
           geom_boxplot()
print(my_plot)


```


The results evidence a significant lower phosphorylation signal on the Y7F mutant. Highlighting the importance of this 7 intracellular tyrosines for the DCBLD2 phosphorylation driven by PDGFR. 
