---
title: "Homework 5. Matrices, lists, and data frames"
author: "Daniel Penados-Richter"
date: "2024-02-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Assign to the variable n_dims a single random integer between 3 and 10.

```{r}
#Create one random integer number between 3 and 10
n_dims <- as.integer(runif(1,min=3,max=10))
typeof(n_dims)
print(n_dims)
#Creates a new vector between 1 and the random generated number from before
new_vec<-(1:(n_dims^2))
print(new_vec)
#Reshuffles last vector
sample(x=new_vec)
#Create a matrix with the last vect. Is a square matrix, because nrow=n_dims
m<-matrix(data = new_vec, nrow = n_dims)
print(m)

#Transpose with t()
m_matrix<-t(m)
print(m_matrix)

#Sum and mean teh first row
print(sum(m_matrix[1,]))
print(mean(m_matrix[1,]))

#Sum and mean the very last row. 
print(sum(m_matrix[n=n_dims,]))
print(mean(m_matrix[n=n_dims,]))


#Coverts the matrix en eigen values
m_eigen<-eigen(m_matrix, only.values = TRUE)
print(m_eigen)

typeof(m_eigen) #They are a list of variables

```


2. Create a list with the following named elements:
my_matrix, which is a 4 x 4 matrix filled with random uniform values
my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
my_letters, which is a 26-element vector of all the lower-case letters in random order.


```{r}

my_matrix<-matrix(runif(16), nrow = 4)
print(my_matrix)

my_logical<-(runif(100)>.5)
print(my_logical)

.<-letters
my_letters<-sample(.)
print(my_letters)
```

Then, complete the following steps:

create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
use the typeof() function to confirm the underlying data types of each component in this list
combine the underlying elements from the new list into a single atomic vector with the c() function.
what is the data type of this vector?

```{r}

my_list<-list(my_matrix[2,2], my_logical[2], my_letters[2])
print(my_list)
#Typeof to verify which type of object. The [[]] are used to determine the object within the list. 
typeof(my_list[[1]])#It is a double
typeof(my_list[[2]])#It is logical
typeof(my_list[[3]])#it is a character

new_vector<-c(my_list[[1]], my_list[[2]], my_list[[3]])
typeof(new_vector) #Its a character. Because character is the one with higher herarqui in the vecttos. 
print(new_vector) #Note that the number and letter are in "" becouse they are characters. 
```

3. Create a data frame with the two variables (= columns) and 26 cases (= rows) below:
-call the first variable my_unis and fill it with 26 random uniform values from 0 to 10
-call the second variable my_letters and fill it with 26 capital letters in random order.
-for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.
-for the first variable, write a single line of R code to identify which rows have the missing values.
-re-order the entire data frame to arrange the second variable in alphabetical order
-calculate the column mean for the first variable.K

```{r}
#Create the variables
my_unis <- as.integer(runif(26,min=0,max=10))
my_letters <- sample(LETTERS)
#Convert in data frame
my_df <-data.frame(my_unis, my_letters)
#Will create a new data frame replacing the variable my_unis to the same variable replacing four random numbers to NA
my_df<-data.frame(my_unis=replace(my_unis, sample(my_unis, size=4), NA), my_letters)

#Print which values are missing
print(which(is.na(my_df)))

#order my_df by my_letters alphabetical order
my_df2 <- my_df[order(my_df$my_letters),]

#Print the mean of my_unis of the data frame my_df, omitting NA values. 
print(mean(my_df2$my_unis, na.rm = TRUE))
#Not committing NA values. Does not work. :(
print(mean(my_df2$my_unis, na.rm = FALSE))


   



```
