---
title: "Homework 8. dplyr and SQL"
author: "Daniel Penados-Richter"
date: "2024-03-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### This homework assignment focus on data manipulation in R. Complete these problems using the dplyr and tidyverse packages.

1. Examine the structure of the iris data set. How many observations and variables are in the data set?

``` {r}
library(dplyr)
library(tidyr)
data("iris")
glimpse(iris)
```

iris has 5 variables, and 150 observations. 

2. Create a new data frame iris1 that contains only the species ```virginica``` and ```versicolor``` with sepal lengths longer than 6 cm and sepal widths longer than 2.5 cm. How many observations and variables are in the data set?

```{r}
iris1<-filter(iris, Sepal.Width>2.5, Sepal.Length>6)%>%
  filter(Species %in% c("virginica", "versicolor"))

```

3. Now, create a ```iris2``` data frame from ```iris1``` that contains only the columns for ```Species, Sepal.Length, and Sepal.Width```. How many observations and variables are in the data set?

```{r}
iris2<-select(iris1, Species, Sepal.Length, Sepal.Width)
glimpse(iris2)
```

It has 56 obervations. 

4. Create an ```iris3``` data frame from ```iris2``` that orders the observations from largest to smallest sepal length. Show the first 6 rows of this data set.

```{r}
iris3<-arrange(iris2, by = Sepal.Length)
head(iris3)

```

5. Create an ```iris4``` data frame from ```iris3``` that creates a column with a sepal area ```(length * width)``` value for each observation. How many observations and variables are in the data set?

```{r}
iris4<-mutate(iris3, Sepal.Area = Sepal.Length*Sepal.Width)
glimpse(iris4)
```

It has 4 variables, 56 observations. 


6. Create ```iris5``` that calculates the average sepal length, the average sepal width, and the sample size of the entire ```iris4``` data frame and print ```iris5```.

```{R}
iris5<-summarize(iris4,  meanLenght = mean(Sepal.Length), 
                 meanWidth = mean(Sepal.Width), 
                 sample.Size = n())
 print(iris5)

```

7. Finally, create iris6 that calculates the average sepal length, the average sepal width, and the sample size for each species of in the iris4 data frame and print iris6.

```{r}
iris_species<-group_by(iris4, Species)
iris6<-summarize(iris_species,  meanLenght = mean(Sepal.Length), 
                 meanWidth = mean(Sepal.Width), 
                 sample.Size = n())
print(iris6)
```
8. In these exercises, you have successively modified different versions of the data frame ```iris1 iris2 iris3 iris4 iris5 iris6```. At each stage, the output data frame from one operation serves as the input fro the next. A more efficient way to do this is to use the pipe operator %>% from the tidyr package. See if you can rework all of your previous statements (except for iris5) into an extended piping operation that uses iris as the input and generates irisFinal as the output.

```{r}
irisFinal<-iris%>%
  filter(Sepal.Width>2.5, Sepal.Length>6)%>%
  filter(Species %in% c("virginica", "versicolor"))%>%
  select(Species, Sepal.Length, Sepal.Width)%>%
  arrange(by = Sepal.Length)%>%
  group_by(Species)%>%
  summarize(meanLenght = mean(Sepal.Length), 
                 meanWidth = mean(Sepal.Width), 
                 sample.Size = n())
print(irisFinal)
```
9. Create a ‘longer’ data frame using the original iris data set with three columns named “Species”, “Measure”, “Value”. The column “Species” will retain the species names of the data set. The column “Measure” will include whether the value corresponds to Sepal.Length, Sepal.Width, Petal.Length, or Petal.Width and the column “Value” will include the numerical values of those measurements.

```{r}
iris_long<-pivot_longer(iris, cols=1:4, names_to = "Measure", values_to = "Values")
glimpse(iris_long)

```
Now it has 3 variables, and 600 obervations. 